### 2019-01-10 code-oneclock-paper/latex/CloTT/Structure/ClockContexts.tex
Work on types/context, added unicode symbols forr syntactic sugar of sigma

\AgdaHide{
\begin{code}%
\>[0]\AgdaKeyword{module}\AgdaSpace{}%
\AgdaModule{CloTT.Structure.ClockContexts}\AgdaSpace{}%
\AgdaKeyword{where}\<%
\end{code}
}

Our goal is to make an interpretation of the syntax as defined in \Cref{sec:syntax}.
For this, we first need to interpret clock contexts and for those there are two possibilities.
It is either empty or we have a single clock.
In the first case, we use the tag \AIC{set} to indicate we interpret types and contexts as sets.
In the second case, we use \AIC{tot}, which says we work in some topos of trees.

\begin{code}%
\>[0]\AgdaKeyword{data}\AgdaSpace{}%
\AgdaDatatype{tag}\AgdaSpace{}%
\AgdaSymbol{:}\AgdaSpace{}%
\AgdaPrimitiveType{Set}\AgdaSpace{}%
\AgdaKeyword{where}\<%
\\
\>[0][@{}l@{\AgdaIndent{0}}]%
\>[2]\AgdaInductiveConstructor{set}\AgdaSpace{}%
\AgdaSymbol{:}\AgdaSpace{}%
\AgdaDatatype{tag}\<%
\\
%
\>[2]\AgdaInductiveConstructor{tot}\AgdaSpace{}%
\AgdaSymbol{:}\AgdaSpace{}%
\AgdaDatatype{tag}\<%
\end{code}